Exploring Conformational Dynamics and Related Time Scales via Nonequilibrium Molecular Dynamics Simulation

### 2013-08-19 writedown/npt-neq/alanine.tex
backup

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
Here we fastly review the approach of performing dynamical nonequilibrium
molecular dynamics (D-NEMD) that was initiated by Giovanni Ciccoti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
We denote the macroscopic nonequilibrium observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
where $\hat O (\vect x)$ is the microscopic observable, 
which is measured at the phase space position $\vect x$,
comparing with the macroscopic observable that is not phase space independent.
In this work, we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known, and is equal to the equilibrium distribution
without any electric field.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes the inner produce in the
phase space.  We assume the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix, then the Liouville equation of the probability
distribution is
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
The Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} O(\vect x, 0)$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the initial
probability distribution is the equilibrium distribution, an
explicit translation of Eq.~\eqref{eqn:tmp4} is that the
nonequilibrium observable is equal to the ensemble average of
microscopic observable measured along trajectory $\vect x(t)$, the
initial configuration of which is subject to the \emph{equilibrium}
distribution. In practice, we firstly run an equilibrium MD simulation
to generate a sampling of configurations that are subject to the
equilibrium distribution. And then by using these configurations as
initial configuration, the Hamiltonian dynamics are integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper. Finally, the macroscopic observable is
estimated by averaging the microscopic observable measured at the end
points of the trajectories.

Being more generalized, if the microscopic observable depends not only on the end
configuration at time $t$ but also depends on the historical
configurations along the trajectory, the expression~\eqref{eqn:tmp4} can be generalize to
\begin{align}
  O(t) = \int d\vect x\,\rho(\vect x, 0) \int_{\mathcal C\{\vect x, 0; t\}} \hat O[\vect x_s] \,d\mathcal P[\vect x_s] 
\end{align}
where the microscopic observable $ \hat O[\vect x_s] $ is now a
functional of the trajectory.  $\mathcal C\{\vect x, 0; t\}$ is the
set of all continuous trajectories starting at point $\vect x$ at time
0, and ending at time $t$. $\mathcal P[\vect x_s] $ is the measure of
the trajectory space $\mathcal C$.  In our case, it is the
$\delta$-function peaked around the trajectories governed by the
Hamiltonian dynamics. To be general, it can also be the measure of all
trajectories generated by, for example, a Langevin dynamics, then the
branching trajectories should be generated by the Langevin
dynamics rather than the Hamiltonian dynamics.

@@

The aforementioned D-NEMD simulation algorithm
does not describe how the boundary of the system should be set up.
The very naive idea of doing the simulation
is to simulate a infinitely large isolated system, or any large enough
isolated subsystem of it.
In practice, it is both impossible and unnecessary to
do such a simulation, due to the unaffordable computational expense.
Therefore, proper boundary condition should be supplied
to truncated the simulation region. It is only possible when we
have a clear knowledge on
on how the nonequilibrium system is set up,
and what are the nonequilibrium observables to measure.

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig/thermostat.eps}
  \caption{Schematic plot of the truncated nonequilibrium subsystem.
    The white part is the Hamiltonian dynamics subsystem
    and the gray part is thermostating region. The whole system
    is subject to the periodic boundary condition.}
  \label{fig:tmp2}
\end{figure}

Since we are mainly interested in the
nonthermal~\cite{delaHoz2005microwaves} effect of the electric field
to the configuration of a solvated alanine dipeptide, we assume the
alanine is embeded into an infinitely large solvent environment, and
the electric field is only applied to the neighborhood of the alanine.
Any extra heat generated by the electric field can be effectively
absorbed by the solvent environment.  It is argued that such an
accurate reaction temperature control is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
Now the question comes to how to truncate the solvent environment
so that the system can be handled by the limited computer resource.
% In practice, it is impossible to simulate an infinitely large
% environment.
For an equilibrium simulation, the solution is to couple
the system to a thermostat, under which the canonical ensemble can be
sampled by only simulating a finite size system with periodic boundary
condition. For a nonequilibrium simulation, simply couple the system
to a thermostat may not be a good idea, because the \emph{dynamics}
generated by the thermostat is artificial, and deviation from the
Hamiltonian dynamics under investigation will introduce artificial
effects in the nonequilibrium observations. To solve this problem, we
observe that only preserving the Hamiltonian dynamics of the alanine 
itself and the water molecules in the nearby solvation shell is
crucial to the interested nonequilibrium observable
(in this work, it is the conformational change of alanine),
while the detailed dynamics of
the water molecules far away is of less importance. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition. The simulating region is divided into two
subregions: Near the alanine dipeptide, the dynamics is kept to be
Hamiltonian, and this subregion is called \emph{the Hamiltonian dynamics
region}. While far from the alanine, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostating
region} (see Fig.~\ref{fig:tmp2}).  For simplicity, the Hamiltonian
dynamics region is assumed to be spherical, centered at the
alpha-carbon of the alanine, and with radius of $R_{ex}$.  Further, we
freeze the movement of the alpha-carbon, so that the alanine is always
located at the center of the simulation region.
Therefore, the dynamics is preseved in the Hamiltonian dynamics region (where the properties of interest are observed)
and the artificial effect of thermostating is negligible.
At the same time, the thermostating region works as a infinitely large
environment that effectively absorbs the extra heat, so that
the nonthermal effect can be studied.
The validity of the above
statement will be later check latter by numerical examples showing that the
nonequilibrium observables do not depends on the size of the system
and the size of the Hamiltonian dynamics subregion, if they are
reasonably large and the finite size effect does not play a dominating role.

@@

\section{Example I: Alanine dipeptide
  under a constant electric field}

@@

In this section,
we want to study the nonequilibrium properties of an alanine dipeptide
under a constant EF.  At time $t=0$~ps, the system
has been fully equilibriated without any electric field. From $t=0$ to
$t=t_{\textrm{init}}$, the eclectic field is switched on linearly, while
after $t=t_{\textrm{init}}$, the field is kept constant at
$E_{\infty}$.  In this work we consider $t_{\textrm{init}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of this field is arbitrarily chosen: along the
$x$ direction. Since the dipole of the equilibrium configurations of
alanine can be along any direction, applying the electric along $x$-axis is
not biased.
Being more general, in this paper we denote the
electric field as a function of time $\vect E(t)$.
Therefore, the constant electric field is
$\vect E(t) = (E_\infty\cdot t/t_{init},0,0)$ as $0\leq t < t_{init}$, and 
$\vect E(t) = (E_\infty,0,0)$
as $t \geq t_{init}$.
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
The electric field exserts force $\vect F = q \vect E(t)$ on
the partial charge $q$ in the system.
The size of the  Hamiltonian dynamics region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
Firstly, an equilibrium NVT simulation of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
50~\textsf{ps}, so we use 2000 initial configurations for each nonequilibrium
MD simulation, if not stated otherwise.
% They server as starting point of the branching
% trajectories.
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostating region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an obviously effect on the
dynamics of the system.

@@

\begin{figure}
  \centering
  \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/c-2.eps}}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-begin-1.eps}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-end-1.eps}
  \caption{The population of metastable states of the alanine dipeptide.
    (a): the
    equilibrium state at $t=0$~\textsf{ps}. (b): The new equilibrium
    state at $t=1000$~\textsf{ps}, i.e., fully relaxed under the
    constant electric field 1~V/nm. The darker color
    indicates larger population on the Ramachandran histogram.}
  \label{fig:tmp4}
\end{figure}

@@

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
metastable state changes due to the external electric field, for
example, the population of state $C$ vanishes in equilibrium,
but plays an important role when the electric field is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the electric field. 
Firstly, we calculate the probability of being in a certain state that
is defined by:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
please notice that the time dependency of the probability, so it is
a nonequilibrium observable.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Secondly,
we study the net probability flux from metastable state $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
which presents the rate of how much conformation $J$ is changed
into $I$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Now project the original nonequilibrium process $\vect x_t$
onto the discretized states $\{A_1, A_2, B_1, B_2, C\}$,
if the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need firstly approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
If one wants a precise finite difference approximation, a smaller $\Delta t$ is
preferable,
but then the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not too small so that the joint probabilities can be calculated properly.
In the oscillating EF case, the net probability flux is also highly oscillating,
so it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which presents the time averaged effect of the EF to the molecular conformation.

@@

The time-dependent probability of the system in a certain metastable state is given in
Fig.~\ref{fig:tmp5}. The beta sheet conformation,
which presents at the beginning,
fades away as the system relaxes to the new equilibrium. The
probability of alpha helix $A_1$ conformation grows from 45\% to 70\%.
Alpha helix $A_2$  does not change a lot under the electric field.
The left-hand helix configuration $C$ noticeably grows from 0\% to 20\%.

@@

% To further test the dynamics of metastable population, two simulations
% of $t_{\textrm{init}} = 100~\textsf{ps}$ and $t_{\textrm{init}} =
% 500~\textsf{ps}$ are performed. The first time scale of $t =
% 20~\textsf{ps}$ disappears in both of the testing simulations. In the
% simulation of $t_{\textrm{init}} = 100~\textsf{ps}$, the population of
% $A$, $B_1$ and $C$ converges to the non-equilibrium steady state at
% around $500~\textsf{ps}$. This further verifies internal time scale of
% $A_1$, $B_1$ and $C$ presented in the testing case of
% $t_{\textrm{init}} = 10~\textsf{ps}$.  The dynamics of $B_2$ is
% faster: it decays to zero as soon as the external electric field is fully
% switched on. In the case of $t_{\textrm{init}} = 500~\textsf{ps}$, the
% growth of the external electric field is comparable to the longest
% internal time scale of system, which means that the changes of the
% perturbation to the system is so slow that the system can catch up, so
% the process can be treated as  quasi-equilibrium.
% The numerical results show the same story: the relaxation
% of all populations are around $500 \sim 600~\textsf{ps}$.

>>>>

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
Here we fastly review the approach of performing dynamical nonequilibrium
molecular dynamics (D-NEMD) that was initiated by Giovanni Ciccoti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
We denote the macroscopic nonequilibrium observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
where $\hat O (\vect x)$ is the microscopic observable, 
which is measured at the phase space position $\vect x$,
comparing with the macroscopic observable that is not phase space independent.
In this work, we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known, and is identical to the equilibrium distribution
without any EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes the inner produce in the
phase space.  We assume the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix, then the Liouville equation of the probability
distribution is
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} \hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the system starts from the equilibrium
distribution, an
explicit translation of Eq.~\eqref{eqn:tmp4} is that the
nonequilibrium observable is equal to the ensemble average of
microscopic observable measured along trajectories, the
initial configurations of which are subject to the \emph{equilibrium}
distribution. In practice, we firstly run an equilibrium MD simulation
to generate a sampling of configurations that are subject to the
equilibrium distribution. And then by using these configurations as
initial configuration, the Hamiltonian dynamics are integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper. Finally, the macroscopic observable is
estimated by averaging the microscopic observable measured at the end
points of the trajectories.

Being more generalized, if the microscopic observable depends not only on the end
configuration at time $t$, but also on the historical
configurations along the trajectory, then expression~\eqref{eqn:tmp4} takes the generalized form:
\begin{align}
  O(t) = \int d\vect x\,\rho(\vect x, 0) \int_{\mathcal C\{\vect x, 0; t\}} \hat O[\vect x_s] \,d\mathcal P[\vect x_s] 
\end{align}
where the microscopic observable $ \hat O[\vect x_s] $ is now a
functional of the trajectory.  $\mathcal C\{\vect x, 0; t\}$ is the
set of all continuous trajectories starting at point $\vect x$ at time
0, and ending at time $t$. $\mathcal P[\vect x_s] $ is the measure of
the trajectory space $\mathcal C$.  In our case, it is the
$\delta$-function peaked around the trajectories governed by the
Hamiltonian dynamics. To be general, it can also be the measure of all
trajectories generated by, for example, a Langevin dynamics, then the
branching trajectories should be generated by the Langevin
dynamics rather than the Hamiltonian dynamics.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig/thermostat.eps}
  \caption{Schematic plot of the nonequilibrium temperature control.
    The white part is the Hamiltonian dynamics region
    and the gray part is thermostating region.
    $R_{ex}$ denotes the size of the Hamiltonian dynamics region, which
    is a sphere located at the center of the box in the Figure.
    The whole system
    is subject to the periodic boundary condition. }
  \label{fig:tmp2}
\end{figure}

% The aforementioned D-NEMD simulation algorithm
% does not describe how the boundary of the system should be set up.
% The very naive idea of doing the simulation
% is to simulate a infinitely large isolated system, or any large enough
% isolated subsystem of it.
% In practice, 
% Therefore, we are interested in a

We are mainly interested in the
nonthermal~\cite{delaHoz2005microwaves} effect of the EF
to the configuration of a solvated alanine dipeptide,
so the temperature control in a nonequilibrium system
becomes an important issue, becuase it was argued that  an
accurate internal reaction temperature control is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
A naive (but close to realistic) implementation of the temperature control is to 
embeded the alanine into an infinitely large solvent environment, and 
the EF is only applied to the neighborhood of the alanine, so that
any extra heat generated by the EF can be effectively
absorbed by the solvent environment.
However, in practice, it is sometime too expensive to explicitly
simulate a very large solvent environment
due to the usually limited computational resources.
Therefore, we are interested in how to cheaply replace the infinitely large
solvent environment by a small affordable system, while
at the same time accurately reproduce the nonequilibrium
physical properties of interest.
% In practice, it is impossible to simulate an infinitely large
% environment.
If only the equilibrium properties are interested, the solution is to couple
the system to a thermostat, under which the desired ensemble
(canonical in this case) can be
sampled by only simulating a finite size system with periodic boundary
condition.
However, directly coupling a thermostat to the nonequilibrium system
will disturb the original Hamiltonian dynamics. This is
in general problematic, 
because the dynamics  under thermostating is artificial designed,
so that the equilibrium distribution can be correctly produce, but not
the nonequilibrium properties of system.
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% However, for a nonequilibrium system, the \emph{dynamics} properties
% will be distrubed by the thermostat, because the
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% For a nonequilibrium simulation, simply couple the system
% to a thermostat may not be a good idea, because the \emph{dynamics}
% generated by the thermostat is artificial, and deviation from the
% Hamiltonian dynamics under investigation will introduce artificial
% effects in the nonequilibrium observations.
To solve this problem, we
observe that only preserving the Hamiltonian dynamics of the alanine 
itself and the water molecules in the nearby solvation shell is
crucial to the interested nonequilibrium observable
(in this work, it is the conformational change of alanine),
while the detailed dynamics of
the water molecules far away is of less importance. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition. Accordingly, the simulating region is divided into two
subregions: Near the alanine dipeptide, the dynamics is kept to be
Hamiltonian, and this subregion is called \emph{the Hamiltonian dynamics
region}. While far from the alanine, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostating
region} (see Fig.~\ref{fig:tmp2}).  For simplicity, the Hamiltonian
dynamics region is assumed to be spherical, centered at the
alpha-carbon of the alanine, and with radius of $R_{ex}$.  Further, we
freeze the movement of the alpha-carbon, so that the alanine is always
located at the center of the simulation region.
Therefore, the dynamics is preseved in the Hamiltonian dynamics region (where the properties of interest are observed)
and the artificial effect of thermostating is negligible due to the finite corrlation
range of liquid water.
At the same time, the thermostating region works as a infinitely large
environment that effectively absorbs the extra heat, so that
the nonthermal effect can be studied.
The validity of the above
statement will be later check latter by numerical examples showing that the
nonequilibrium observables do not depends on the size of the system
and the size of the Hamiltonian dynamics subregion, if they are
reasonably large and the finite size effect does not play a dominating role.

@@

\section{Example I: Alanine dipeptide
  under a constant EF}

@@

In this section,
we want to study the nonequilibrium properties of an alanine dipeptide
under a constant EF.  At time $t=0$~ps, the system
has been fully equilibriated without any EF. From $t=0$ to
$t=t_{\textrm{init}}$, the eclectic field is switched on linearly, while
after $t=t_{\textrm{init}}$, the field is kept constant at
$E_{\infty}$.  In this work we consider $t_{\textrm{init}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of this field is arbitrarily chosen: along the
$x$ direction. Since the dipole of the equilibrium configurations of
alanine can be along any direction, applying the electric along $x$-axis is
not biased.
Being more general, in this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, the constant EF is
$\vect E(t) = (E_\infty\cdot t/t_{init},0,0)$ as $0\leq t < t_{init}$, and 
$\vect E(t) = (E_\infty,0,0)$
as $t \geq t_{init}$.
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and solved in 644 TIP3P
water molecules.
The EF exserts force $\vect F = q \vect E(t)$ on
the partial charge $q$ in the system.
The size of the  Hamiltonian dynamics region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
Firstly, an equilibrium NVT simulation of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
50~\textsf{ps}, so we use 2000 initial configurations for each nonequilibrium
MD simulation, if not stated otherwise.
% They server as starting point of the branching
% trajectories.
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostating region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
The whole system is also coupled to a Parrinello-Rahman barostat with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at the ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an obviously effect on the
dynamics of the system.

@@

\begin{figure}
  \centering
  \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/c-2.eps}}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-begin-1.eps}
  \includegraphics[width=0.37\textwidth]{fig.ala/ext.mode1.010.Ex.01.00.t1000ps.recheck/fig-end-1.eps}
  \caption{The population of metastable states of the alanine dipeptide.
    (a): the
    equilibrium state at $t=0$~\textsf{ps}. (b): The new equilibrium
    state at $t=1000$~\textsf{ps}, i.e., fully relaxed under the
    constant EF 1~V/nm. The darker color
    indicates larger population on the Ramachandran histogram.}
  \label{fig:tmp4}
\end{figure}

@@

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
metastable state changes due to the external EF, for
example, the population of state $C$ vanishes in equilibrium,
but plays an important role when the EF is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the EF. 
Firstly, we calculate the probability of being in a certain state that
is defined by:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
please notice that the time dependency of the probability, so it is
a nonequilibrium observable.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Secondly,
we study the net probability flux from metastable state $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
which presents the rate of how much conformation $J$ is changed
into $I$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Now project the original nonequilibrium process $\vect x_t$
onto the discretized states $\{A_1, A_2, B_1, B_2, C\}$,
if the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need firstly approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
If one wants a precise finite difference approximation, a smaller $\Delta t$ is
preferable,
but then the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not too small so that the joint probabilities can be calculated properly.
In the oscillating EF case, the net probability flux is also highly oscillating,
so it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which presents the time averaged effect of the EF to the molecular conformation.

@@

The time-dependent probability of the system in a certain metastable state is given in
Fig.~\ref{fig:tmp5}. The beta sheet conformation,
which presents at the beginning,
fades away as the system relaxes to the new equilibrium. The
probability of alpha helix $A_1$ conformation grows from 45\% to 70\%.
Alpha helix $A_2$  does not change a lot under the EF.
The left-hand helix configuration $C$ noticeably grows from 0\% to 20\%.

@@

% To further test the dynamics of metastable population, two simulations
% of $t_{\textrm{init}} = 100~\textsf{ps}$ and $t_{\textrm{init}} =
% 500~\textsf{ps}$ are performed. The first time scale of $t =
% 20~\textsf{ps}$ disappears in both of the testing simulations. In the
% simulation of $t_{\textrm{init}} = 100~\textsf{ps}$, the population of
% $A$, $B_1$ and $C$ converges to the non-equilibrium steady state at
% around $500~\textsf{ps}$. This further verifies internal time scale of
% $A_1$, $B_1$ and $C$ presented in the testing case of
% $t_{\textrm{init}} = 10~\textsf{ps}$.  The dynamics of $B_2$ is
% faster: it decays to zero as soon as the external EF is fully
% switched on. In the case of $t_{\textrm{init}} = 500~\textsf{ps}$, the
% growth of the external EF is comparable to the longest
% internal time scale of system, which means that the changes of the
% perturbation to the system is so slow that the system can catch up, so
% the process can be treated as  quasi-equilibrium.
% The numerical results show the same story: the relaxation
% of all populations are around $500 \sim 600~\textsf{ps}$.

\begin{figure}
  \centering
  % \includegraphics[width=0.3\textwidth]{fig.ala/ele.field/field.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/field.dipole/fig-dipol-order-conf.eps}
  % \\
  % \includegraphics[width=0.18\textwidth]{fig.ala/confs/ext.mode1.010.Ex.01.00.t1000ps.recheck/a1-1.eps}
  % \includegraphics[width=0.18\textwidth]{fig.ala/confs/ext.mode1.010.Ex.01.00.t1000ps.recheck/a2-1.eps}
  % \includegraphics[width=0.18\textwidth]{fig.ala/confs/ext.mode1.010.Ex.01.00.t1000ps.recheck/b1-1.eps}
  % \includegraphics[width=0.18\textwidth]{fig.ala/confs/ext.mode1.010.Ex.01.00.t1000ps.recheck/b2-1.eps}
  % \includegraphics[width=0.18\textwidth]{fig.ala/confs/ext.mode1.010.Ex.01.00.t1000ps.recheck/c-1.eps}
  \caption{The
    $x$-component of the dipole moment, and the orientational order parameter
    of the alanine molecule as a
    function of time. The dashed line presents the strengths of the
    external EF, while the solid line presents the dipole
    moments. The left vertical axis is for the dipole moment, while
    the right is for the orientational order parameter.
  }
  \label{fig:tmp3}
\end{figure}
